I owe an enormous debt of gratitude to many people. I am particularly grateful to my dissertation advisor, David Siegel. I am incredibly fortunate to have found an advisor as deeply committed to mentorship and excited by the possibilities of research as he is.  I am not alone in finding his faith in the abilities of his students to be profoundly empowering.  Without his encouragement, guidance, and patience, this project would exist in a much lesser form. 

I am also grateful to my committee members, Laia Balcells, Peter Feaver, Livia Schubiger, and Jacob Shapiro, for their guidance and valuable contributions.  Katherine Heller's advice, perspective, and friendship were invaluable for the development of this dissertation. Similarly, Kyle Beardsley and David Banks provided significant insight at critical stages of the project.

The dissertation has benefited tremendously from generous comments and suggestions from many scholars, particularly Amelia Hoover-Green, Kyle Endres, Mia Bloom,  Rocio Titiunik,  Sherry Zacks, Ora Szekely, and Victor Asal; participant feedback from the 2019 Online Peace Science Colloquium; the 2020 New Faces Conference; and panel and attendee comments at the American Political Science Association, International Studies Association, Peace Science Society, and Visions in Methodology annual meetings. I benefited tremendously from comments and constructive criticism from members of the political science departments at Florida State University, Rice University, the University of North Carolina at Chapel Hill, the University of Pennsylvania, Pennsylvania State University, and Princeton University. The project was sharpened through engagement with visitors to the Duke University Political Science Workshop Speakers, through the Carnegie IPSCON program, and the participants and organizers of the Organizations and Their Effectiveness summer institute at the Center for Advanced Study in the Behavioral Sciences.  Finally, support from the Bradley Foundation, the Evans Fund, the Department of Political Science, and The Graduate School at Duke University made the research possible.

I would like to thank my colleagues at the SITE Intelligence Group, particularly Rita Katz, Adam Raisman, and Sam Monk, for the years of friendship and their intellectual companionship and generosity. More than almost anyone else in the world, the three of you have shaped who I am as an adult. As well, I appreciate the support of many friends and colleagues at Duke, especially Katie Webster, Juan Tellez, Jeremy Spater, Shahryar Minhas,  Hao Liu, Marco Morucci, So Jin Lee, Gloria Cheung, and the women of Duke Women in Political Science.

Finally, I am indebted to my parents and sisters for their encouragement and adventurousness and to Drew for being the best and most supportive partner. Without you, I would be rudderless.


